---
title: "Russian military losses in Ukraine"
output: github_document
date: "December 2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

#### The replication analysis for A.Bessudnov. (2022). "Ethnic and regional inequalities in the Russian military fatalities in the 2022 war in Ukraine". SocArXiv. https://osf.io/preprints/socarxiv/s43yf . DOI 10.31235/osf.io/s43yf .



```{r}
library(tidyverse)
library(geofacet)
library(DescTools)
library(Hmisc)
library(ggpubr)
library(lubridate)

options(tibble.width = Inf)

df <- read_csv("dfCasualties.csv")
```

## Fatalities over time

```{r}

df |>
  filter(deathConfirmedDate < dmy("2.12.2022")) |>
  ggplot(aes(x = deathConfirmedDate)) +
  geom_histogram(bins = 50) +
  xlab("Date when death was confirmed")

```


## Number of men killed and mortality rate per capita by region

```{r}

# Read the 2021 census data.

census2021 <- read_csv("census2021_nMen16-615.csv")

# This file has a variable nMen16_615_2021: the number of men in the working age (16-61.5) by region
# as per census 2021 (2020).

# Number of men killed by region and mortality rate per 100,000 men aged 16-61.5

table_mortRates <- df |>
  filter(!is.na(region)) |>
  count(region) |>
  # join with the census data
  left_join(census2021, by = c("region" = "regNameEng")) |>
  # calculate mortality rate per 100,000 men, with confidence intervals
  mutate(mortalityRate = map2(n, nMen16_615_2021, ~ prop.test(.x, .y, conf.level=0.95) |>
                                     broom::tidy())) |>
  unnest(mortalityRate) |>
  mutate(mortalityRate = estimate * 100000) |>
  mutate(mortalityRate_lowCI = conf.low * 100000) |>
  mutate(mortalityRate_highCI = conf.high * 100000) |>
  arrange(-mortalityRate) |>
  select(region, regNameISO, n, nMen16_615_2021, mortalityRate, mortalityRate_lowCI, mortalityRate_highCI) 

table_mortRates |>
  print(n = Inf)
```

## Map of mortality rates by region

```{r}
# This sets a grid for a map.

ru_federal_subjects_2022 <- data.frame(
  row = c(1, 1, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4, 4, 4, 
          4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 
          5, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 7, 7, 7, 7, 7, 7, 7, 8, 
          8, 8, 9, 9, 9, 9, 9, 9, 9, 10, 10, 10, 10, 11, 11, 11),
  col = c(1, 5, 17, 18, 4, 10, 1, 11, 16, 17, 3, 4, 5, 9, 10, 14, 11, 12, 8, 9,  
          10, 16, 17, 19, 1, 3, 13, 14, 15, 4, 5, 6, 7, 11, 12, 7, 8, 9, 10, 16,
          13, 14, 15, 3, 4, 5, 6, 11, 7, 8, 9, 10, 17, 12, 13, 14, 15, 3, 4, 5, 
          6, 7, 8, 9, 14, 4, 5, 6, 5, 6, 7, 2, 4, 5, 6, 7, 8, 1, 5, 6, 7, 8, 5, 6, 7),
  code = c("SPE", "MUR", "CHU", "KAM", "KR", "NEN", "MOW", "YAN", "SA", 
           "MAG", "LEN", "NGR", "VLG", "ARK", "KO", "KYA", "KHM", "TYU", 
           "ME", "KIR", "PER", "AMU", "KHA", "SAK", "KGD", "PSK", "TOM", 
           "KEM", "IRK", "TVE", "YAR", "IVA", "KOS", "SVE", "KGN", "NIZ", 
           "CU", "TA", "UD", "YEV", "NVS", "KK", "BU", "SMO", "KLU", 
           "MOS", "VLA", "CHE", "MO", "ULY", "SAM", "BA", "PRI", "OMS", 
           "ALT", "TY", "ZAB", "BRY", "ORL", "TUL", "RYA", "PNZ", "SAR", 
           "ORE", "AL", "KRS", "LIP", "TAM", "BEL", "VOR", "VGG", "UA-43", 
           "AD", "KDA", "ROS", "KL", "AST",  "UA-40" ,"KC", "STA", "CE", "DA", 
           "KB", "SE", "IN"),
  name = c("St.Petersburg", "Murmansk", "Chukotka", 
           "Kamchatka", "Karelia", "Nenets", 
           "Moscow", "Yamalo-Nenets", "Sakha", 
           "Magadan", "Leningrad", "Novgorod", 
           "Vologda", "Arkhangelsk", "Komi", 
           "Krasnoyarsk", "Khanty-Mansi", 
           "Tyumen", "Mari El", "Kirov", 
           "Perm", "Amur", "Khabarovsk", 
           "Sakhalin", "Kaliningrad", "Pskov", 
           "Tomsk", "Kemerovo", "Irkutsk", 
           "Tver", "Yaroslavl", "Ivanovo", 
           "Kostroma", "Sverdlovsk", "Kurgan", 
           "Nizhny Novgorod", "Chuvashia", "Tatarstan", 
           "Udmurtia", "Yevreyskaya", 
           "Novosibirsk", "Khakassia", "Buryatia", 
           "Smolensk", "Kaluga", "Moscow Oblast", 
           "Vladimir", "Chelyabinsk", "Mordovia", 
           "Ulyanovsk", "Samara", "Bashkortostan", 
           "Primorsky", "Omsk", "Altai Krai", "Tuva", 
           "Zabaykalsky", "Bryansk", "Oryol", 
           "Tula", "Ryazan", "Penza", 
           "Saratov", "Orenburg", "Altai Republic", 
           "Kursk", "Lipetsk", "Tambov", 
           "Belgorod", "Voronezh", "Volgograd", 
           "Crimea", "Adygea", "Krasnodar", 
           "Rostov", "Kalmykia", "Astrakhan", 
           "Sevastopol",
           "Karachay-Cherkessia", "Stavropol", 
           "Chechnya", "Dagestan", 
           "Kabardino-Balkaria", "North Ossetia", 
           "Ingushetia"),
  stringsAsFactors = FALSE
)

df_map_mortRates <- table_mortRates |>
  select(regNameISO, mortalityRate) |>
  rename(code = regNameISO) |>
  mutate(x = rep(1,85),
         y = rep(1,85))

# Map

map1 <- ggplot(df_map_mortRates, aes(x = x, y = y, fill = log(mortalityRate), label = code)) +
  geom_col() +
  geom_text(x = 1, y = 0.5, size = 4) +
  facet_geo(~ code, grid = ru_federal_subjects_2022) +
  theme_void() +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
  ) +
  labs(fill="Log mortality rate \n(per 100,000 men)") +
  scale_fill_distiller(palette = "Spectral", direction = -1)

map1

```

## Relative risks by ethnic group for Russia

```{r}
# Number and proportion among the dead by ethnic group (assumed from the name)

df |>
  count(impliedEthnicity) |>
  mutate(perc = n / sum(n)) |>
  arrange(-n)

# read in the data from census 2010 on ethnic proportions in the male Russian population
# (Note that people with missing data on ethnicity were excluded from the calculation of proportions)

census2010EthnPropPopulationAll <- read_csv("census2010EthnPropPopulationAll.csv")

# aggregated and compared with the 2010 proportions

df_relRisk <- df |>
  mutate(newEthn = case_when(
    impliedEthnicity == "BelRusUkr" ~ "BelRusUkr",
    impliedEthnicity == "BashTat" ~ "BashTat",
    impliedEthnicity == "CheDagIng" ~ "CheDagIngKabAdKarBalOs",
    impliedEthnicity == "KabAdKarBalOs" ~ "CheDagIngKabAdKarBalOs",
    impliedEthnicity == "Buryat" ~ "Buryat",
    impliedEthnicity == "Tuvan" ~ "Tuvan",
    TRUE ~ "other"
  )) |>
  count(newEthn) |>
  mutate(sumKilled = sum(n)) |>
  mutate(prop = n / sum(n)) |>
  # join in data from the 2010 census
  full_join(census2010EthnPropPopulationAll, by = "newEthn") |>
  mutate(relRisk = prop / propPopulation) |>
  arrange(-relRisk)

df_relRisk


# calculate confidence intervals for the risk ratios with DescTools (RelRisk())

# The data should be presented as a 2x2 matrix.
# In the first row: n killed / (n total in the data - n killed)
# In the second row: n ethnic group (male pop 2010) / (n all men 2010 - n ethnic group male 2010)

# This function calculates relative risk with 95% CIs for an ethnic group.

relRiskCI <- function(x, ethn = ethn) {
    x |>
    filter(newEthn == ethn) |>
    select(n, sumKilled, nPopulation, sumPopulation) |>
    rename(nKilled = n) |>
    mutate(nNotKilled = sumKilled - nKilled) |>
    mutate(nPop_minusGroup = sumPopulation - nPopulation) |>
    select(nKilled, nNotKilled, nPopulation, nPop_minusGroup) |>
    as.numeric() |>
    matrix(nrow = 2, byrow = TRUE) |>
    RelRisk(conf.level = 0.95) 
}

# An example for Tuvans

relRiskCI(df_relRisk, ethn = "Tuvan")
  
# Risk ratios with CIs for all ethnic groups

df_RR_results <- data.frame()
df_RR_results

k <- 1
for (i in df_relRisk$newEthn) {
  df_RR_results <- rbind(df_RR_results, c(i, relRiskCI(df_relRisk, ethn = i)))
  k <- k + 1
}
colnames(df_RR_results) <- c("ethn", "relRisk", "lowRelRiskCI", "highRelRiskCI")

df_RR_results

```

## For some ethnic republics: proportion ethnic Russian names dead vs proportion of ethnic non-Russian names dead

Republics to explore: Buryatia, Tuva, North Ossetia, Dagestan, Chechnya, Ingushetia, Adygeya, Bashkortostan, Kalmykia, Kabardino-Balkaria, Tatarstan, Karachaevo-Chekessiya, Astrakhan.

```{r}

# Read in the 2010 census data with the data on the ethnic composition by region

census_2010_ethn <- read_csv("census_2010_ethn.csv")

df_relRisk_region <- df |>
  filter(region %in% c("Buryatia", "Tuva", "North Ossetia",
                       "Dagestan", "Astrakhan.", "Chechnya",
                       "Adygea", "Ingushetia", "Bashkortostan",
                       "Kalmykia", "Kabardino-Balkaria", "Tatarstan",
                       "Karachay-Cherkessia")) |>
  mutate(ethnRusUkrBel = case_when(
    impliedEthnicity == "BelRusUkr" ~ 1,
    TRUE ~ 0
  )) |>
  mutate(nonethnRusUkrBel = case_when(
    impliedEthnicity == "BelRusUkr" ~ 0,
    TRUE ~ 1
  )) |>
  group_by(region) |>
  summarise(
    nKilled = n(),
    nEthnRusUkrBel_killed = sum(ethnRusUkrBel, na.rm = TRUE),
    nNonEthnRusUkrBel_killed = sum(nonethnRusUkrBel, na.rm = TRUE)
  ) |>
  # join with the census data
  left_join(census_2010_ethn, by = c("region" = "regNameEng")) |>
  select(region, nKilled, nNonEthnRusUkrBel_killed, nMen_ethn_nonmissing, nRusUkrBelMen) |>
  mutate(propNonRusDead = nNonEthnRusUkrBel_killed / nKilled) |>
  mutate(propNonRusAll = (nMen_ethn_nonmissing - nRusUkrBelMen) / nMen_ethn_nonmissing) |>
  mutate(rr = propNonRusDead / propNonRusAll)

df_relRisk_region |>
  arrange(-rr) |>
  print(n = Inf)

# a function to calculate relative risks by region with confidence intervals
  
relRiskCI_region <- function(x, regName = region) {
    x |>
    filter(region == regName) |>
    select(nKilled, nNonEthnRusUkrBel_killed, nMen_ethn_nonmissing, nRusUkrBelMen) |>
    mutate(nRusKilled = nKilled - nNonEthnRusUkrBel_killed) |>
    mutate(nNonRusPop = nMen_ethn_nonmissing - nRusUkrBelMen) |>
    select(nNonEthnRusUkrBel_killed, nRusKilled, nNonRusPop, nRusUkrBelMen) |>
    as.numeric() |>
    matrix(nrow = 2, byrow = TRUE) |>
    RelRisk(conf.level = 0.95) 
}

# An example for Tuva.

relRiskCI_region(df_relRisk_region, regName = "Tuva")
  
# Risk ratios with CIs for all regions

df_RR_results_region <- data.frame()
df_RR_results_region

k <- 1
for (i in df_relRisk_region$region) {
  df_RR_results_region <- rbind(df_RR_results_region, c(i, relRiskCI_region(df_relRisk_region, regName = i)))
  k <- k + 1
}
colnames(df_RR_results_region) <- c("region", "relRisk", "lowRelRiskCI", "highRelRiskCI")

df_RR_results_region

```

The relative risk shows if ethnic minorities are more likely to die than ethnic Russians. If RR > 1 ethnic minority names are more common among the dead, if RR < 1 the ethnic minority names are less common.

## Correlations with regional characteristics

```{r}

# Read in regional characteristics for 2020:
# percent below the subsistence level
# unemployment rate
# subsistence level (4th quarter 2020)
# mean income per capita

regionData <- read_csv("regionCharacteristics.csv")

# Joining in with the data on regional mortality rates.

regionData_joined <- table_mortRates |>
  select(-regNameISO) |>
  left_join(regionData, by = c("region" = "regNameEng")) |>
  # calculate the regional income / subsistence level ratio
  mutate(incomeToSubsitenceRatio = meanIncome / subsistenceLevel) |>
  select(-regNameISO) |>
  # Join in the data on proportion non-Russian population.
  left_join(census_2010_ethn, by = c("region" = "regNameEng")) |>
  mutate(percentNonRus = 100 - percRus)

```

```{r}
# Scatter plots with non-parametric smooth

# Percent non-ethnically Russian populationvs. mortality rate

regionData_joined |>
  ggplot(aes(x = percentNonRus, y = mortalityRate)) +
  geom_point() +
  geom_smooth() +
  xlab("Percent non-ethnically Russian population") +
  ylab("Regional mortality rate")

# Data for percent non-ethnically Russian for Crimea and Sevastopol are not available from the 2010 census; these regions have been dropped.

# Percent with the incomes below the subsistence level

regionData_joined |>
  ggplot(aes(x = percBelowSubsistence, y = mortalityRate)) +
  geom_point() +
  geom_smooth() +
  xlab("Percent with income below subsistence") +
  ylab("Regional mortality rate")


# Correlation matrix

regionData_joined |>
  select(mortalityRate, percentNonRus, percBelowSubsistence, unemploymentRate, incomeToSubsitenceRatio) |>
  as.matrix() |>
  rcorr()

# Regression model: mortality rate ~ percent below subsistence level + percent non-Russian

m1 <- lm(mortalityRate ~ percBelowSubsistence + percentNonRus, regionData_joined)

summary(m1)


# Plots: mortality rate vs percent below subsistence level and percent non-ethnically Russian

plot1 <- regionData_joined |>
  ggplot(aes(x = percBelowSubsistence, y = mortalityRate, label = regNameISO)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(nudge_x = 1) +
  stat_cor(aes(label = ..r.label..), size = 8) +
  xlab("Percent of incomes below subsistence level (2020)") +
  ylab("Mortality rate per 100,000 men") +
  theme(axis.title=element_text(size=16))


plot2 <- regionData_joined |>
  ggplot(aes(x = percentNonRus, y = mortalityRate, label = regNameISO)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(nudge_x = 3) +
  stat_cor(aes(label = ..r.label..), size = 8) +
  xlab("Percent of non-ethnically Russian population (2010)") +
  ylab("Mortality rate per 100,000 men") +
  theme(axis.title=element_text(size=16))


figure2 <- ggarrange(plot1, plot2)
figure2

```

## Prepare an anonymised data set for Github

```{r}

df_anon <- df |>
  mutate(firstName = "removed") |>
  mutate(lastName = "removed") |>
  select(-c(impliedEthnicity, regNameISO))

write_csv(df_anon, "dfCasualtiesAnonymised.csv", na = "")

```
