---
title: "Russian military fatalities in Ukraine"
output: github_document
date: "April 2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

#### The replication analysis for A.Bessudnov. (2023). "Ethnic and regional inequalities in the Russian military fatalities in Ukraine: Preliminary findings from crowdsourced data". SocArXiv. https://osf.io/preprints/socarxiv/s43yf . DOI 10.31235/osf.io/s43yf .



```{r}
library(tidyverse)
library(geofacet)
library(DescTools)
library(Hmisc)
library(ggpubr)
library(lubridate)
library(zoo)

options(tibble.width = Inf)

df <- read_csv("dfFatalities12April2023.csv")
```

## Total number of fatalities in the data set

```{r}
df |>
  count()
```

## Fatalities over time

```{r}

df |>
  mutate(monthDeath = as.yearmon(deathConfirmedDate)) |>
  count(monthDeath) |>
  ggplot(aes(x = monthDeath, y = n)) +
  geom_col() +
  xlab("Month when death was confirmed")

```

## Number of men killed and mortality rate per capita by region

```{r}

# Read the 2021 census data.

census2021 <- read_csv("census2021_nMen16-615.csv")

# This file has a variable nMen16_615_2021: the number of men in the working age (16-61.5) by region
# as per census 2021 (2020).

# Number of men killed by region and mortality rate per 100,000 men aged 16-61.5

table_mortRates <- df |>
  filter(!is.na(region)) |>
  count(region) |>
  # join with the census data
  left_join(census2021, by = c("region" = "regNameEng")) |>
  # calculate mortality rate per 100,000 men, with confidence intervals
  mutate(mortalityRate = map2(n, nMen16_615_2021, ~ prop.test(.x, .y, conf.level=0.95) |>
                                     broom::tidy())) |>
  unnest(mortalityRate) |>
  mutate(mortalityRate = estimate * 100000) |>
  mutate(mortalityRate_lowCI = conf.low * 100000) |>
  mutate(mortalityRate_highCI = conf.high * 100000) |>
  arrange(-mortalityRate) |>
  select(region, regNameISO, n, nMen16_615_2021, mortalityRate, mortalityRate_lowCI, mortalityRate_highCI) 

table_mortRates |>
  print(n = Inf)
```

## Map of mortality rates by region

```{r}
# This sets a grid for a map.

ru_federal_subjects_2022 <- data.frame(
  row = c(1, 1, 2, 2, 2, 2, 2, 3, 3, 3, 3, 3, 3, 3, 3, 3, 4, 4, 4, 4, 4, 4, 4, 
          4, 4, 4, 4, 4, 4, 4, 4, 4, 4, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 5, 
          5, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 6, 7, 7, 7, 7, 7, 7, 7, 8, 
          8, 8, 9, 9, 9, 9, 9, 9, 9, 10, 10, 10, 10, 11, 11, 11),
  col = c(1, 5, 17, 18, 4, 10, 1, 11, 16, 17, 3, 4, 5, 9, 10, 14, 11, 12, 8, 9,  
          10, 16, 17, 19, 1, 3, 13, 14, 15, 4, 5, 6, 7, 11, 12, 7, 8, 9, 10, 16,
          13, 14, 15, 3, 4, 5, 6, 11, 7, 8, 9, 10, 17, 12, 13, 14, 15, 3, 4, 5, 
          6, 7, 8, 9, 14, 4, 5, 6, 5, 6, 7, 2, 4, 5, 6, 7, 8, 1, 5, 6, 7, 8, 5, 6, 7),
  code = c("SPE", "MUR", "CHU", "KAM", "KR", "NEN", "MOW", "YAN", "SA", 
           "MAG", "LEN", "NGR", "VLG", "ARK", "KO", "KYA", "KHM", "TYU", 
           "ME", "KIR", "PER", "AMU", "KHA", "SAK", "KGD", "PSK", "TOM", 
           "KEM", "IRK", "TVE", "YAR", "IVA", "KOS", "SVE", "KGN", "NIZ", 
           "CU", "TA", "UD", "YEV", "NVS", "KK", "BU", "SMO", "KLU", 
           "MOS", "VLA", "CHE", "MO", "ULY", "SAM", "BA", "PRI", "OMS", 
           "ALT", "TY", "ZAB", "BRY", "ORL", "TUL", "RYA", "PNZ", "SAR", 
           "ORE", "AL", "KRS", "LIP", "TAM", "BEL", "VOR", "VGG", "UA-43", 
           "AD", "KDA", "ROS", "KL", "AST",  "UA-40" ,"KC", "STA", "CE", "DA", 
           "KB", "SE", "IN"),
  name = c("St.Petersburg", "Murmansk", "Chukotka", 
           "Kamchatka", "Karelia", "Nenets", 
           "Moscow", "Yamalo-Nenets", "Sakha", 
           "Magadan", "Leningrad", "Novgorod", 
           "Vologda", "Arkhangelsk", "Komi", 
           "Krasnoyarsk", "Khanty-Mansi", 
           "Tyumen", "Mari El", "Kirov", 
           "Perm", "Amur", "Khabarovsk", 
           "Sakhalin", "Kaliningrad", "Pskov", 
           "Tomsk", "Kemerovo", "Irkutsk", 
           "Tver", "Yaroslavl", "Ivanovo", 
           "Kostroma", "Sverdlovsk", "Kurgan", 
           "Nizhny Novgorod", "Chuvashia", "Tatarstan", 
           "Udmurtia", "Yevreyskaya", 
           "Novosibirsk", "Khakassia", "Buryatia", 
           "Smolensk", "Kaluga", "Moscow Oblast", 
           "Vladimir", "Chelyabinsk", "Mordovia", 
           "Ulyanovsk", "Samara", "Bashkortostan", 
           "Primorsky", "Omsk", "Altai Krai", "Tuva", 
           "Zabaykalsky", "Bryansk", "Oryol", 
           "Tula", "Ryazan", "Penza", 
           "Saratov", "Orenburg", "Altai Republic", 
           "Kursk", "Lipetsk", "Tambov", 
           "Belgorod", "Voronezh", "Volgograd", 
           "Crimea", "Adygea", "Krasnodar", 
           "Rostov", "Kalmykia", "Astrakhan", 
           "Sevastopol",
           "Karachay-Cherkessia", "Stavropol", 
           "Chechnya", "Dagestan", 
           "Kabardino-Balkaria", "North Ossetia", 
           "Ingushetia"),
  stringsAsFactors = FALSE
)

df_map_mortRates <- table_mortRates |>
  select(regNameISO, mortalityRate) |>
  rename(code = regNameISO) |>
  mutate(x = rep(1,85),
         y = rep(1,85))

# Map

map1 <- ggplot(df_map_mortRates, aes(x = x, y = y, fill = log(mortalityRate), label = code)) +
  geom_col() +
  geom_text(x = 1, y = 0.5, size = 4) +
  facet_geo(~ code, grid = ru_federal_subjects_2022) +
  theme_void() +
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
  ) +
  labs(fill="Log mortality rate \n(per 100,000 men)") +
  scale_fill_distiller(palette = "Spectral", direction = -1)

map1

```

## Relative risks by ethnic group for Russia

```{r}
# Number and proportion among the dead by ethnic group (assumed from the name)

df |>
  count(impliedEthnicity) |>
  mutate(perc = n / sum(n)) |>
  arrange(-n)

# read in the data from census 2010 on ethnic proportions in the male Russian population
# (aged 8 to 32 at the time of the 2010 census, corresponding to 20 to 44 in 2022)
# (Note that people with missing data on ethnicity were excluded from the calculation of proportions)

census2010EthnPropPopulationAll <- read_csv("census2010EthnPropPopulationAll.csv")

# aggregated and compared with the 2010 proportions

df_relRisk <- df |>
  mutate(newEthn = case_when(
    impliedEthnicity == "BelRusUkr" ~ "BelRusUkr",
    impliedEthnicity == "BashTat" ~ "BashTat",
    impliedEthnicity == "CheDagIng" ~ "CheDagIngKabAdKarBalOs",
    impliedEthnicity == "KabAdKarBalOs" ~ "CheDagIngKabAdKarBalOs",
    impliedEthnicity == "Buryat" ~ "Buryat",
    impliedEthnicity == "Tuvan" ~ "Tuvan",
    TRUE ~ "other"
  )) |>
  count(newEthn) |>
  mutate(sumKilled = sum(n)) |>
  mutate(prop = n / sum(n)) |>
  # join in data from the 2010 census
  full_join(census2010EthnPropPopulationAll, by = "newEthn") |>
  mutate(relRisk = prop / propPopulation) |>
  mutate(relRisk_4_50 = prop / propPopulation_4_50) |>
  mutate(relRisk_8_32 = prop / propPopulation_8_32) |>
  arrange(-relRisk_8_32)


df_relRisk


# calculate confidence intervals for the risk ratios with DescTools (RelRisk())

# The data should be presented as a 2x2 matrix.
# In the first row: n killed / (n total in the data - n killed)
# In the second row: n ethnic group (male pop 2010) / (n all men 2010 - n ethnic group male 2010)

# This function calculates relative risk with 95% CIs for an ethnic group.
relRiskCI <- function(x, ethn = ethn) {
    x |>
    filter(newEthn == ethn) |>
    select(n, sumKilled, nPopulation_8_32, sumPopulation_8_32) |>
    rename(nKilled = n) |>
    mutate(nNotKilled = sumKilled - nKilled) |>
    mutate(nPop_minusGroup_8_32 = sumPopulation_8_32 - nPopulation_8_32) |>
    select(nKilled, nNotKilled, nPopulation_8_32, nPop_minusGroup_8_32) |>
    as.numeric() |>
    matrix(nrow = 2, byrow = TRUE) |>
    RelRisk(conf.level = 0.95) 
}

# An example for Tuvans

relRiskCI(df_relRisk, ethn = "Tuvan")
  
# Risk ratios with CIs for all ethnic groups

df_RR_results <- data.frame()
df_RR_results

k <- 1
for (i in df_relRisk$newEthn) {
  df_RR_results <- rbind(df_RR_results, c(i, relRiskCI(df_relRisk, ethn = i)))
  k <- k + 1
}
colnames(df_RR_results) <- c("ethn", "relRisk", "lowRelRiskCI", "highRelRiskCI")

df_RR_results
```

## For some ethnic republics: proportion ethnic Russian names dead vs proportion of ethnic non-Russian names dead

Republics to explore: Buryatia, Tuva, North Ossetia, Dagestan, Chechnya, Ingushetia, Adygeya, Bashkortostan, Kalmykia, Kabardino-Balkaria, Tatarstan, Karachaevo-Chekessiya, Astrakhan.

Risk ratios calculated with the data on male population aged 8 to 32 in 2010 (20 to 44 in 2022).

```{r}

# Read in the 2010 census data with the data on the ethnic composition by region

census_2010_ethn <- read_csv("census_2010_ethn.csv")

df_relRisk_region <- df |>
  filter(region %in% c("Buryatia", "Tuva", "North Ossetia",
                       "Dagestan", "Astrakhan", "Chechnya",
                       "Adygea", "Ingushetia", "Bashkortostan",
                       "Kalmykia", "Kabardino-Balkaria", "Tatarstan",
                       "Karachay-Cherkessia")) |>
  mutate(ethnRusUkrBel = case_when(
    impliedEthnicity == "BelRusUkr" ~ 1,
    TRUE ~ 0
  )) |>
  mutate(nonethnRusUkrBel = case_when(
    impliedEthnicity == "BelRusUkr" ~ 0,
    TRUE ~ 1
  )) |>
  group_by(region) |>
  summarise(
    nKilled = n(),
    nEthnRusUkrBel_killed = sum(ethnRusUkrBel, na.rm = TRUE),
    nNonEthnRusUkrBel_killed = sum(nonethnRusUkrBel, na.rm = TRUE)
  ) |>
  # join with the census data
  left_join(census_2010_ethn, by = c("region" = "regNameEng")) |>
  select(region, nKilled, nNonEthnRusUkrBel_killed, nMen_ethn_nonmissing, nMen_ethn_nonmissing_4_50, nMen_ethn_nonmissing_8_32, nRusUkrBelMen, nRusUkrBelMen_4_50, nRusUkrBelMen_8_32) |>
  mutate(propNonRusDead = nNonEthnRusUkrBel_killed / nKilled) |>
  mutate(propNonRusAll = (nMen_ethn_nonmissing - nRusUkrBelMen) / nMen_ethn_nonmissing) |>
  mutate(propNonRusAll_4_50 = (nMen_ethn_nonmissing_4_50 - nRusUkrBelMen_4_50) / nMen_ethn_nonmissing_4_50) |>
  mutate(propNonRusAll_8_32 = (nMen_ethn_nonmissing_8_32 - nRusUkrBelMen_8_32) / nMen_ethn_nonmissing_8_32) |>
  mutate(rr = propNonRusDead / propNonRusAll) |>
  mutate(rr_4_50 = propNonRusDead / propNonRusAll_4_50) |>
  mutate(rr_8_32 = propNonRusDead / propNonRusAll_8_32)


df_relRisk_region |>
  arrange(-rr_8_32) |>
  print(n = Inf)

# a function to calculate relative risks by region with confidence intervals
  
relRiskCI_region <- function(x, regName = region) {
    x |>
    filter(region == regName) |>
    select(nKilled, nNonEthnRusUkrBel_killed, nMen_ethn_nonmissing_8_32, nRusUkrBelMen_8_32) |>
    mutate(nRusUkrBelKilled = nKilled - nNonEthnRusUkrBel_killed) |>
    mutate(nNonRusPop_8_32 = nMen_ethn_nonmissing_8_32 - nRusUkrBelMen_8_32) |>
    select(nNonEthnRusUkrBel_killed, nRusUkrBelKilled, nNonRusPop_8_32, nRusUkrBelMen_8_32) |>
    as.numeric() |>
    matrix(nrow = 2, byrow = TRUE) |>
    RelRisk(conf.level = 0.95) 
}

# An example for Tuva.

relRiskCI_region(df_relRisk_region, regName = "Tuva")
  
# Risk ratios with CIs for all regions

df_RR_results_region <- data.frame()
df_RR_results_region

k <- 1
for (i in df_relRisk_region$region) {
  df_RR_results_region <- rbind(df_RR_results_region, c(i, relRiskCI_region(df_relRisk_region, regName = i)))
  k <- k + 1
}
colnames(df_RR_results_region) <- c("region", "relRisk", "lowRelRiskCI", "highRelRiskCI")

df_RR_results_region

```

The relative risk shows if ethnic minorities are more likely to die than ethnic Russians. If RR > 1 ethnic minority names are more common among the dead, if RR < 1 the ethnic minority names are less common.

## Distribution by age

```{r}
# mean age
df |>
  summarise(
    mean(age, na.rm = TRUE)
  )

# distribution by age

df |>
  ggplot(aes(x = age)) +
  geom_bar() +
  xlab("Age") +
  ylab("Number")

```

## Number of deaths and proportion Slavic by type of military force

```{r}
df |>
  mutate(BelRusUkr = case_when(
      impliedEthnicity == "BelRusUkr" ~ 1,
      TRUE ~ 0
    )) |>
  group_by(unitType) |>
  summarise(
    nType = n(),
    propSlavic = mean(BelRusUkr, na.rm = TRUE)
    ) |>
  mutate(propType = nType / sum(nType)) |>
  select(unitType, nType, propType, propSlavic)


```

## Deaths by month and type of military unit

```{r}
unitType12Apr <- df |>
  mutate(monthDeath = as.yearmon(deathConfirmedDate)) |>
  mutate(unitType = case_when(
    is.na(unitType) ~ "missing data",
    TRUE ~ unitType
  )) |>
  mutate(unitType = factor(unitType,
            levels = c("inmates", "private military company", "conscripted", "volunteers",
                       "police / national guard / FSB", "professional army", "missing data"))) |>
  filter(!is.na(monthDeath)) |>
  count(unitType, monthDeath) |>
  ggplot(aes(x = monthDeath, y = n, fill = unitType)) +
  geom_col() +
  ylab("Number of deaths") +
  xlab("Date of death confirmation") +
  # theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  labs(fill = "Unit type") +
  theme(text = element_text(size=20))

unitType12Apr

```

## Changes over time by ethnic group and region

```{r}
ethnTrend12Apr <- df |>
  mutate(newEthn = case_when(
    impliedEthnicity == "BelRusUkr" ~ "Slavic",
    impliedEthnicity == "BashTat" ~ "BashTat",
    impliedEthnicity == "CheDagIng" ~ "North Caucasian",
    impliedEthnicity == "KabAdKarBalOs" ~ "North Caucasian",
    impliedEthnicity == "Buryat" ~ "Buryat",
    impliedEthnicity == "Tuvan" ~ "Tuvan",
    TRUE ~ "other"
  )) |>
  mutate(monthDeath = as.yearmon(deathConfirmedDate)) |>
  count(monthDeath, newEthn) |>
  group_by(monthDeath) |>
  mutate(propMonth = n / sum(n)) |>
  filter(newEthn %in% c("Slavic", "Buryat", "Tuvan", "North Caucasian")) |>
  filter(monthDeath > "Feb 2022") |>
  ggplot(
    aes(x = monthDeath, y = propMonth)
  ) +
  geom_point() +
  geom_smooth(colour = "black", se = FALSE) +
  xlab("Date of death confirmation") +
  ylab("Proportion among deaths") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  facet_wrap(~ newEthn, scales="free_y") +
  theme(text = element_text(size=20))

ethnTrend12Apr 


dfMacroRegionMonth <- df |>
  mutate(macroRegion = case_when(
      region == "Arkhangelsk" ~ "North / East",
      region == "Nenets" ~ "North / East",
      region == "Tyumen" ~ "North / East",
      region == "Khanty-Mansi" ~ "North / East",
      region == "Yamalo-Nenets" ~ "North / East",
      region == "Altai Krai" ~ "North / East",
      region == "Krasnodar" ~ "South",
      region == "Krasnoyarsk" ~ "North / East",
      region == "Primorsky" ~ "North / East",
      region == "Stavropol" ~ "South",
      region == "Khabarovsk" ~ "North / East",
      region == "Amur" ~ "North / East",
      region == "Astrakhan" ~ "South",
      region == "Belgorod" ~ "Central",
      region == "Bryansk" ~ "Central",
      region == "Vladimir" ~ "Central",
      region == "Volgograd" ~ "South",
      region == "Vologda" ~ "Central",
      region == "Voronezh" ~ "Central",
      region == "Nizhny Novgorod" ~ "Volga",
      region == "Ivanovo" ~ "Central",
      region == "Irkutsk" ~ "North / East",
      region == "Ingushetia" ~ "South",
      region == "Kaliningrad" ~ "Central",
      region == "Tver" ~ "Central",
      region == "Kaluga" ~ "Central",
      region == "Kamchatka" ~ "North / East",
      region == "Kemerovo" ~ "North / East",
      region == "Kirov" ~ "Volga",
      region == "Kostroma" ~ "Central",
      region == "Samara" ~ "Volga",
      region == "Kurgan" ~ "North / East",
      region == "Kursk" ~ "Central",
      region == "St.Petersburg" ~ "Central",
      region == "Leningrad" ~ "Central",
      region == "Lipetsk" ~ "Central",
      region == "Magadan" ~ "North / East",
      region == "Moscow" ~ "Central",
      region == "Moscow Oblast" ~ "Central",
      region == "Murmansk" ~ "North / East",
      region == "Novgorod" ~ "Central",
      region == "Novosibirsk" ~ "North / East",
      region == "Omsk" ~ "North / East",
      region == "Orenburg" ~ "Volga",
      region == "Oryol" ~ "Central",
      region == "Penza" ~ "Volga",
      region == "Perm" ~ "Volga",
      region == "Pskov" ~ "Central",
      region == "Rostov" ~ "South",
      region == "Ryazan" ~ "Central",
      region == "Saratov" ~ "Volga",
      region == "Sakhalin" ~ "North / East",
      region == "Sverdlovsk" ~ "North / East",
      region == "Smolensk" ~ "Central",
      region == "Tambov" ~ "Central",
      region == "Tomsk" ~ "North / East",
      region == "Tula" ~ "Central",
      region == "Ulyanovsk" ~ "Volga",
      region == "Chelyabinsk" ~ "North / East",
      region == "Zabaykalsky" ~ "North / East",
      region == "Chukotka" ~ "North / East",
      region == "Yaroslavl" ~ "Central",
      region == "Adygea" ~ "South",
      region == "Bashkortostan" ~ "Volga",
      region == "Buryatia" ~ "North / East",
      region == "Dagestan" ~ "South",
      region == "Kabardino-Balkaria" ~ "South",
      region == "Altai Republic" ~ "North / East",
      region == "Kalmykia" ~ "South",
      region == "Karelia" ~ "North / East",
      region == "Komi" ~ "North / East",
      region == "Mari El" ~ "Volga",
      region == "Mordovia" ~ "Volga",
      region == "North Ossetia" ~ "South",
      region == "Karachay-Cherkessia" ~ "South",
      region == "Tatarstan" ~ "Volga",
      region == "Tuva" ~ "North / East",
      region == "Udmurtia" ~ "Volga",
      region == "Khakassia" ~ "North / East",
      region == "Chechnya" ~ "South",
      region == "Chuvashia" ~ "Volga",
      region == "Sakha" ~ "North / East",
      region == "Yevreyskaya" ~ "North / East",
      region == "Sevastopol" ~ "South",
      region == "Crimea" ~ "South")
  ) |>
  mutate(monthDeath = as.yearmon(deathConfirmedDate)) |>
  count(monthDeath, macroRegion) |>
  group_by(monthDeath) |>
  mutate(propDeaths = n / sum(n)) 
      
dfMacroRegionMonth 

regionTrend12Apr <- dfMacroRegionMonth  |>
   filter(monthDeath > "Feb 2022") |>
   filter(!is.na(macroRegion)) |>
   ggplot(aes(x = monthDeath, y = propDeaths)) +
   geom_point() +
   geom_smooth(colour = "black", se = FALSE) +
  xlab("Date of death confirmation") +
  ylab("Proportion among deaths") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  facet_wrap(~ macroRegion)  +
  theme(text = element_text(size=20))

regionTrend12Apr
```

## Correlations with regional characteristics

```{r}

# Read in regional characteristics for 2020:
# percent below the subsistence level
# unemployment rate
# subsistence level (4th quarter 2020)
# mean income per capita

regionData <- read_csv("regionCharacteristics.csv")

# Joining in with the data on regional mortality rates.

regionData_joined <- table_mortRates |>
  select(-regNameISO) |>
  left_join(regionData, by = c("region" = "regNameEng")) |>
  # calculate the regional income / subsistence level ratio
  mutate(incomeToSubsitenceRatio = meanIncome / subsistenceLevel) |>
  select(-regNameISO) |>
  # Join in the data on proportion non-Russian population.
  left_join(census_2010_ethn, by = c("region" = "regNameEng")) |>
  mutate(percentNonRus = 100 - percRus)

```

```{r}
# Scatter plots with non-parametric smooth

# Percent non-ethnically Russian populationvs. mortality rate

regionData_joined |>
  ggplot(aes(x = percentNonRus, y = mortalityRate)) +
  geom_point() +
  geom_smooth() +
  xlab("Percent non-ethnically Russian population") +
  ylab("Regional mortality rate")

# Data for percent non-ethnically Russian for Crimea and Sevastopol are not available from the 2010 census; these regions have been dropped.

# Percent with the incomes below the subsistence level

regionData_joined |>
  ggplot(aes(x = percBelowSubsistence, y = mortalityRate)) +
  geom_point() +
  geom_smooth() +
  xlab("Percent with income below subsistence") +
  ylab("Regional mortality rate")



# Regression model: mortality rate ~ percent below subsistence level + percent non-Russian

m1 <- lm(mortalityRate ~ percBelowSubsistence + percentNonRus, regionData_joined)

summary(m1)


# Plots: mortality rate vs percent below subsistence level and percent non-ethnically Russian

plot1 <- regionData_joined |>
  ggplot(aes(x = percBelowSubsistence, y = mortalityRate, label = regNameISO)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(nudge_x = 1) +
  stat_cor(aes(label = ..r.label..), size = 8) +
  xlab("Percent of incomes below subsistence level (2020)") +
  ylab("Mortality rate per 100,000 men") +
  theme(axis.title=element_text(size=16))


plot2 <- regionData_joined |>
  ggplot(aes(x = percentNonRus, y = mortalityRate, label = regNameISO)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(nudge_x = 3) +
  stat_cor(aes(label = ..r.label..), size = 8) +
  xlab("Percent of non-ethnically Russian population (2010)") +
  ylab("Mortality rate per 100,000 men") +
  theme(axis.title=element_text(size=16))


figure2 <- ggarrange(plot1, plot2)
figure2

```

## Validation with the Yandex data

I collected data on Yandex searches by region (https://wordstat.yandex.com/#!/regions) for two search expressions: "сво выплаты" ("special military operation payments", over 25,000 monthly searches) and "выплаты погибшим" ("payments for killed", over 50,000 monthly searches). The data were collected on 10 February 2023. These are not historical data so it reflects the situation as in January 2023.

(Update: in another round of data collection, I collected Yandex data on 19 April.)

The data show regional popularity of search term, i.e. the affinity index which is the regional the share of searches for the word/expression divided over the total share of all searches in that region.

The correlation between the Yandex searches and the fatality rate is 0.4 or 0.5, depending on the search term.

```{r}

# Reading Yandex searches data

yandex_df <- read_csv("yandex_searches.csv")

# Joining with the fatality rates data

yandex_df <- yandex_df |>
  left_join(table_mortRates, by = "regNameISO")

# "svo vyplaty" (10 Feb)

cor(yandex_df$svo_vyplaty_10Feb, yandex_df$mortalityRate)

# "vyplaty pogibshim" (10 Feb)

cor(yandex_df$vyplaty_pogibshim_10Feb, yandex_df$mortalityRate)

# April 2023 data

# vyplaty pogibshim (19 Apr)
cor(yandex_df$vyplaty_pogibshim_19Apr, yandex_df$mortalityRate)



yandex_df |>
  ggplot(aes(x = vyplaty_pogibshim_10Feb, y = mortalityRate, label = regNameISO)) +
  geom_smooth() +
  geom_text() +
  xlab("vyplaty pogibshim search regional affinity index (10 Feb)") +
  ylab("Fatality rate")

yandex_df |>
  ggplot(aes(x = vyplaty_pogibshim_19Apr, y = mortalityRate, label = regNameISO)) +
  geom_smooth() +
  geom_text() +
  xlab("vyplaty pogibshim search regional affinity index (19 Apr)") +
  ylab("Fatality rate")

```

## Proportion of records with archived source

```{r}
df |>
  summarise(
    sum(is.na(source))
  )

1355 / 20253

```

## Prepare an anonymised data set for Github

```{r}

df_anon <- df |>
  mutate(first_name = "removed") |>
  mutate(last_name = "removed") |>
  mutate(impliedEthnicity = "removed")

write_csv(df_anon, "dfFatalitiesAnonymised.csv", na = "")

```
